---
title: "Plot-to-Map: Tool for comparing biomass plots and maps"
author: "Arnan Araza, PhD candidate, Wageningen University and Research"
date: "October 15, 2020"
project: "ESA-CCI Biomass project / PhD topic"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
## ------------------ Preliminaries ------------------
rm(list=ls())

# packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(rgdal,rgeos,raster,plyr,dplyr,foreach,purrr,BIOMASS,data.table,
               parallel,doParallel,plotrix,gfcanalysis,sf,stringr, randomForest,BIOMASS)

# global variables
mainDir <- "D:/BiomassCCI_2019"
scriptsDir <- "D:/BiomassCCI_2019/scripts" 
outDir <- "D:/BiomassCCI_2019/results"
dataDir <- "D:/BiomassCCI_2019/data"
#plotsFile <- 'SamplePlots.csv'
#plotsFile1 <- 'SamplePoly.csv'
#plotsFile1 <- 'PolyTropiSAR.csv'
agbTilesDir <- "E:/CCIBiomass" #*
treeCoverDir <- 'E:/treecover2010_v3' #*
SRS <- CRS('+init=epsg:4326')
flDir <- 'E:/GFCFolder' 
forestTHs <- 10 
mapYear <- 19
AGBown <- 'NA'
plots <- 'NA'
    #* be sure to download/access tiles

# functions
setwd(scriptsDir)  
source('Polygonize.R')
source('Deforested.R')
source('BiomePair.R')
source('TempFix.R')
source('TempVis.R')
source('MakeBlockPolygon.R')
source('TileNames.R')
source('BlockMeans.R')
source('invDasymetry.R')
source('Plots.R')
source('Nested.R')
source('MeasurementErr.R')
source('StrataAGB.R')
setwd(mainDir)

knitr::opts_chunk$set(echo = TRUE)
```

## Background
Aboveground biomass (AGB) maps provide data and information of the carbon stored by the terrestrial vegetation. AGB maps are made possible by remote sensing (RS) as satellite signals from vegetations can be translated to biomass. Have a look at this new global map used mainly for climate and carbon modelling: 

```{r AGB map}
agb_glob <- raster(paste0(dataDir,'/glob_25km.tif'))
plot(agb_glob, main='GlobBiomass 2010 (Mg/ha)')
```

Aside from global applications, AGB maps are generated to have carbon numbers for national reporting i.e. REDD+. 

```{r echo=FALSE}
ph <- raster('D:/Internship_BAKCUP/data/rf_2018-06-02 04_33_15.tif')
plot(ph, main='AGB map (Mg/ha), Philippines')
```

But these maps are imperfect so they should be validated using "independent" in-situ/ground-truth data i.e. biomass plot data. But like the maps, plot data are not error-free. Have a look at this boxplot of plot uncertainty when using small plots below 0.3ha for validating a global map of 1km resolution. 

```{r}
sd_samp <- read.csv(paste0(dataDir,'/sd_plt.csv'))
boxplot(sd_samp$sd, main='uncertainty (Mg/ha) of small plots when compared with AGB map (1km)')
```

## Functionalities
So before comparing plots and maps, their uncertainties should be accounted! That is the main function of the Plot-to-Map tool (p2m). The largest plot uncertainty comes from tree measurement because biomass is indirectly measured (Chave et al. 2004). So while estimating the biomass of every tree, p2m can also estimate the uncertainty from tree measurement errors. We adopted the error propagation method of BIOMASS R pakcage (Rejou-Mechain et al. 2017) where errors from tree diameter, height, wood density, and allometric model parameters are all propagated.

For plot data with tree-level measurements, diameter is required and if wood density and tree height are missing, those variables can be obtained from a regional/global database and Height-Diameter model, respectively. Take a look at the portion of this sample data:

```{r message=FALSE ,warning=FALSE}
plotTree<- read.csv(paste0(dataDir, '/SampleTree.csv')) 
xyTree <- read.csv(paste0(dataDir,'/SampleTreeXY.csv'))
plotTree$id <- factor(plotTree$id, levels=unique(plotTree$id),labels=seq_along(nrow(plotTree)))
xyTree$id <- factor(xyTree$id, levels=unique(xyTree$id), labels=seq_along(nrow(xyTree)))
head(plotTree)
```

To estimate the uncertianty from measurement errors, we need to use the function "MeasurementErr.R" which is a wrapper of BIOMASS and will further scale the plot biomass (and uncertainty) values per hectare. 

```{r }
plots <- MeasurementErr(plotTree, xyTree, 'World')
head(plots)
```

If tree-level data is unavailable, an alternative to estimate the same uncertainty is to use a pre-trained model that "predicts" the measurement error based on the AGB, plot size, and eco-region/biome. The function "BiomePair" will automatically label the corresponding biomes (e.g. Temperate Coniferous Forest) of your plots.  

```{r message=FALSE, warning=FALSE}
# get biomes and zones
plots <- read.csv(paste0(dataDir,'/SamplePlots.csv'))
plots <- plots[sample(nrow(plots), nrow(plots)*(1 - 0.8)),] #sample a few for faster demo
plots <- BiomePair(plots) #biome labeller
plotsPred <- plots[,c('AGB_T_HA','SIZE_HA', 'GEZ')]
names(plotsPred) <- c('agb', 'size', 'gez')
plotsPred$size <- plotsPred$size * 10000 #convert size to m2
plotsPred$size <- as.integer(plotsPred$size)
plotsPred$gez = factor(plotsPred$gez,
                         levels = c("Boreal","Subtropical","Temperate","Tropical"))
rf <- get(load(paste0(dataDir, '/rf1.RData'))[1]) #pre-trained RF model from 8000+ plots across all major biomes 
plots$sdTree <- predict(rf, plotsPred)
head(plots)
plots0 <- plots 
```

In the case of Wales forest inventory, a plot ("square") is discretized by strata ("sections") and circular plots are established per strata making the dataset a nested plot. The dataset comes from a database (.mdb) and the key to use it is to extract the "NFI trees" using the RODBC function in R. The centroid coordinates of the circular plots are also needed. The function "Nested" will format the tree data and estimate tree biomass from those "alive" trees. For missing wood density, the database for Europe is used. 

```{r message=FALSE, warning=FALSE}
cent <- readOGR(dsn = dataDir, layer = "SampleCentroid") #Wales sample data
tree <- read.csv(paste0(dataDir,'/SampleTreeNested.csv'))
TreeData <- Nested(cent, tree) #disabled the interactive option, not working with when .Rmd is knitted
plotTree <- TreeData[[1]]
xyTree <- TreeData[[2]]
plotTree$id <- factor(plotTree$id, levels=unique(plotTree$id), labels=seq_along(nrow(plotTree)))
xyTree$id <- factor(xyTree$id, levels=unique(xyTree$id), labels=seq_along(nrow(xyTree)))
print(plotTree[40:50,]) #print 20 rows
plots <- MeasurementErr(plotTree, xyTree, 'Europe') #AGB and measurement error
plots
```
To estimate the plot-level ("square") biomass, a weighted mean based on strata size is proposed. The function "StrataAGB" will do this. 

```{r}
StrataAGB(plots)
```


Next, your plot biomass data may need a bit of preprocessing if will be used for AGB map validation (and also calibration/mapping). First, if the plot data is surveyed before or after the AGB map epoch, the function "TempApply" and "TempVar" will modify the plot biomass based on growth rate data unique to specific biomes and continents. Functions "HistoShift" and "HistoVis" summarizes the outcomes.

```{r}
# apply growth data to whole plot data by identifying AGB map year
plots <- plots0
gez <- sort(as.vector((unique(plots$GEZ)))) #get unique gez and without NA (sorting removes it also)
plots.tf <- ldply(lapply (1:length(gez), function(x) 
  TempApply(plots, gez[[x]], 2017)), data.frame) #change the year!
#tree growth data uncertainty estimate
plots.var <- ldply(lapply (1:length(gez), function(x) 
  TempVar(plots, gez[[x]], 2017)), data.frame) 
#get absolute uncertainty of temporally adjusted plots 
plots.tf$sdGrowth <- abs(plots.tf$AGB_T_HA - plots.var$SD)
#order pre and post temproal fix plots for pairing
plots1 <- plots[with(plots, order(GEZ)), ]
plots.tf$AGB_T_HA_ORIG <- plots1$AGB_T_HA
#histogram of temporal fix effect
HistoTemp(plots.tf, 2017)
HistoShift(plots.tf, 2017)
rm(plots1, plots, plots.var) 

```

Another preprocessing step is to scale the plot biomass based on the "forest" variation between the AGB map pixel and plot. This used the global forest change (Hansel et al. 2013) dataset to identify the forest fraction of the map pixel and multiply it to the plot biomass (function "InvDasymetry"). The GFC dataset is also to remove already deforested plots relative to the AGB map epoch (function "Deforested"). 

Lastly, weighted comparison of plot data with the AGB map is needed to account for plot uncertainty. The inverse of the plot variance is used to estimate the weighted mean. This is applicable for validating global maps where the comparison is aggregated at coarser scales i.e. 0.1 degree resolution. The comparison results can be visualized and tabulated: 

## Results
```{r}
plots01 <- get(load(paste0(dataDir,'/GLOB_2010.Rdata')))
Binned(plots01$plotAGB_10, plots01$mapAGB, 'GlobBiomass 2010 plot-to-map comparison', 'samp_p2m.png') 
plotsSLB <- get(load(paste0(dataDir,'/SLB_2017.Rdata')))
Binned(plotsSLB$plotAGB_10, plotsSLB$mapAGB, 'SLB 2017 plot-to-map comparison', 'samp_slb.png') 
```

Thank you and for more info, you can access Plot-to-map with technical documentation at: 

### More info:
+ Araza et al., (2020). Comparing aboveground biomass from plots and global maps: Towards a comprehensive uncertainty assessment. Under preparation. 
+ https://github.com/arnanaraza/PlotToMap for Shiny App
+ https://github.com/arnanaraza/BiomassCCI for Local use

### References: 
+ Chave, J., Condit, R., Aguilar, S., Hernandez, A., Lao, S., & Perez, R. (2004). Error propagation and scaling for tropical forest biomass estimates. Philosophical Transactions of the Royal Society of London. Series B: Biological Sciences, 359(1443), 409-420.
+ Réjou-Méchain, M., Tanguy, A., Piponiot, C., Chave, J., & Hérault, B. (2017). biomass: An r package for estimating above-ground biomass and its uncertainty in tropical forests. Methods in Ecology and Evolution, 8(9), 1163-1167.


